---
  html_document:
    toc: true
    highlight: zenburn
    theme: united
---
The code for this report is on github [here](https://github.com/hbc/rubin-cell-fate-conversion).

```{r setup, echo=FALSE}
knitr::opts_chunk$set(tidy=TRUE, highlight=TRUE, dev="png",
               cache=TRUE, highlight=TRUE, autodep=TRUE, warning=FALSE, error=FALSE,
               message=FALSE, prompt=TRUE, comment='', fig.cap='')
```
```{r load-libaries}
library(useful)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(stringr)
```

Read in the data for each plate into one big dataframe, make unique IDs for
each batch/well combination and use those IDs as the column names.

```{r read-plates, eval=!file.exists("plates.RData")}
setwd("~/rubin-cell-fate-conversion")

# Creating a function to organize the data
plate_reader = function(fn) {
  batch = strsplit(basename(fn), ".", fixed=TRUE)[[1]][1]
  data = read.table(fn, header=TRUE, row.names=1)
  colnames(data) = paste(batch, colnames(data), sep="_")
  data
}
plates_fns = sort(list.files("data", pattern="\\.dat", full.names=TRUE))
plates = do.call(cbind, lapply(plates_fns, plate_reader))
save(plates, file="plates.RData")
```

```{r load-plates}
load("plates.RData")
```

Create a dataframe of the metadata about each sample. This has an identifier
for a sample, Which well it came from, which batch and what it was treated with.

```{r massage-well-identity-data}
welldata_fn = "metadata/Compound Layout 384w1B.csv"
welldata = read_csv(welldata_fn) %>%
  gather(column, treatment, -row) %>%
  mutate(well=paste(row, column, sep="")) %>%
  dplyr::select(well, treatment)
identities = data.frame(str_split_fixed(colnames(plates), "_", 3))
colnames(identities) = c("batch", "drop", "well")
identities$id = colnames(plates)
welldata = identities %>% dplyr::select(batch, well) %>%
         left_join(welldata, by="well")
rownames(welldata) = colnames(plates)
welldata$sample = colnames(plates)
```


Verify that the samples match up between the read counts and the metadata dataframe
and make sure there are no NA counts.

```{r verify-data}
dim(welldata)
dim(plates)
table(rownames(welldata) %in% colnames(plates))
corner(plates)
corner(welldata)
table(complete.cases(plates))
```

Looks like we are good to go. While we're at it we will load the positive
control data.

```{r load-positive-control-data}
positive_fn = "data/Feo_positive_controls.unq.refseq.umi.SC_and_Myoblast_Raw.txt"
positive = read.table(positive_fn, header=TRUE, row.names=rownames(plates))
positive$id = NULL
positive_samples = data.frame(str_split_fixed(colnames(positive), "_", 3))
identities = data.frame(str_split_fixed(colnames(plates), "_", 3))$X1
positive_welldata = data.frame(batch=positive_samples$X1,
                               treatment=positive_samples$X1,
                               well=positive_samples$X1)
```

Now we will calculate some summary statistics about each sample.

```{r plot-genes-detected}
welldata$genes_detected = colSums(plates > 0)
welldata$genes_detected_zscore = ave(welldata$genes_detected, FUN=scale)
welldata$genes_detected_pval = 2 * pnorm(-abs(welldata$genes_detected_zscore))
welldata$genes_detected_padj = p.adjust(welldata$genes_detected_pval, method="BH")
outliers = subset(welldata, genes_detected_padj < 0.1)
ggplot(outliers, aes(treatment, genes_detected, fill=batch)) +
   geom_bar(stat='identity', position='dodge') +
   facet_wrap(~batch) +
   geom_hline(yintercept=ave(welldata$genes_detected), linetype="dashed") +
   ylab("genes with counts > 0") +
   xlab("") +
   theme_bw() +
   scale_x_discrete(breaks=NULL)
```

The M2 plate seems to have lower counts than the plates.

```{r low-counts-plate}
ggplot(welldata, aes(batch, genes_detected)) + geom_boxplot() +
   ylab("genes with counts > 0") +
   xlab("") +
   theme_bw()
```

We can see it in the boxplot.

```{r boxplot-counts-plate}
welldata$counts = colSums(plates)
ggplot(welldata, aes(batch, counts)) + geom_boxplot() +
   ylab("total counts") +
   xlab("") +
   theme_bw()
```

And the number of failed wells, where we call a well failed if it
had < 10,000 total counts.

```{r failed-wells}
welldata$failed = colSums(plates) < 10000
welldata %>% group_by(batch) %>% summarise(nfailed = sum(failed))
```

We can also see in an estimate of library complexity that the M2 library
is less complex than the other libraries.

```{r plot-complexity}
ggplot(welldata, aes(counts, genes_detected, color=batch)) +
   geom_smooth(fill=NA) +
   ylab("genes with counts > 0") +
   xlab("total counts") +
   theme_bw()
```

In this histogram of the genes detected, we can see there are a set of
samples with a low amount of genes detected.

# Render HTML file of .rmd
library(rmarkdown)
render("Sample_comparison_broad_LSP.Rmd")